---
title: "Extending travelpaths"
author: ""
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Extending travelpaths}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  chunk_output_type: console
---


```{r knitr_settings, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", out.width = "100%", error = TRUE)
if (isTRUE(as.logical(Sys.getenv("NO_EXTENDING_VIGNETTE")))) {
  knitr::opts_chunk$set(eval = FALSE)
}
```


```{r, echo = FALSE, message = FALSE}
suppressPackageStartupMessages(library("travelpaths"))
data("tf_mini", package = "trackframe")
```

**travelpaths** uses a plugin system modelled after the
[parsnip](https://parsnip.tidymodels.org/) pattern from tidymodels.
A specification object captures what algorithm to run and with which
parameters.  A separate registry maps each (model, mode, engine) triple to
a concrete fitting function.  The two sides are joined at the point of
calling `fit()`.

The public API for extending the package consists of three functions:

| Function | Purpose |
|---|---|
| `new_travelpath_spec(cls, args, eng_args, mode, engine)` | Create a new specification object |
| `register_fit_function(model, mode, engine, fit_function)` | Register the engine specific fit function |
| `register_argument_mapping(model, mode, engine, argument_mapping)` | Rename arguments before they are passed to the fit function |


# Writing a specification constructor

A specification constructor is a function that collects 
parameters into a named list and passes them to `new_travelpath_spec()`.

```{r}
threshold_stops <- function(
  mode = "segmentation",
  engine = "infostop",
  r1 = 10,
  min_size = 2L,
  min_staying_time = 300L,
  max_time_between = 86400L,
  output_col = "stop_id"
) {
  args <- list(
    r1 = r1,
    min_size = min_size,
    min_staying_time = min_staying_time,
    max_time_between = max_time_between,
    stop_id_col = output_col
  )
  new_travelpath_spec(
    "threshold_stops",
    args = args,
    eng_args = list(),
    mode = mode,
    engine = engine
  )
}
```

`new_travelpath_spec()` validates every argument with
[checkmate](https://mllg.github.io/checkmate/) and returns an object
inheriting from `"travelpath_spec"`:

```{r}
spec <- threshold_stops()
class(spec)
```

```{r}
str(spec)
```

- `args` holds the parameters that are present in all `threshold_stops` implementations. 
- `eng_args` holds parameters that are engine-specific and should not be
exposed in the constructor signature (pass `list()` when unused).


# Registering a fit function

`register_fit_function(model, mode, engine, fit_function)` adds the fit
function to the registry so that `fit()` dispatches correctly based on
the specification object.

```{r eval=FALSE}
register_fit_function(
  "threshold_stops", "segmentation", "infostop", infostop::identify_stops
)
```

The `fit_function` must accept `data` as its **first** positional argument,
followed by keyword arguments matching the names defined in `args` and
`eng_args`.  If the caller passes extra `...` arguments to `fit()`, they
are appended last and override arguments from the specification.


# Updating a specification

`update.travelpath_spec()` lets callers change individual parameters
without rebuilding the spec from scratch.

```{r}
spec |> update(r1 = 25, min_size = 3L) |> str()
```

Use `fresh = TRUE` to reset all unspecified parameters to their constructor
defaults before applying the new values:

```{r}
# Start from threshold_stops(r1 = 99, min_size = 9L), reset everything,
# then apply r1 = 5.
threshold_stops(r1 = 99, min_size = 9L) |>
  update(fresh = TRUE, r1 = 5) |> str()
```

A `params` named list is an alternative to `...` and can be combined with it:

```{r}
update(spec, params = list(r1 = 50), min_size = 4L) |> str()
```


# Registering an argument mapping

Engine functions sometimes use different parameter names than the canonical
travelpaths names. In this case you have two options

1. write a wrapper function that does additional validation and the mapping of the arguments,
2. if there are only small differences you can register an argument mapping.

`register_argument_mapping()` declares a renaming that
is applied automatically inside `fit()` before the call is assembled.

The `argument_mapping` argument is a named character vector where each name
is the spec-level name and the corresponding value is the engine-level name:

```{r eval=FALSE}
register_argument_mapping(
  "change_point_permtest", "segmentation", "cpt",
  c("n_permutations" = "n")
)
```

With this mapping in place, a spec arg called `n_permutations` is
renamed to `n` before being forwarded to `cpt::change_point_test`.


# The fit() dispatch pipeline

`fit.travelpath_spec()` orchestrates the following steps every time a spec
is fitted:

1. Look up the registered fit function for `(model, mode, engine)`.
2. Build the argument list: `c(list(data), args, eng_args)`.
3. Apply any registered argument mappings (see `register_argument_mapping()`).
4. Merge in extra arguments supplied via `...`.
5. Validate the assembled argument list against the fit function signature.
6. Evaluate the call and attach a `<mode>_fit` class to the result.

```{r}
result <- fit(threshold_stops(r1 = 20), tf_mini)
result
```


# Dry-run mode

Passing `dry_run = TRUE` to `fit()` skips evaluation and returns the
assembled call object instead.  This is the recommended way to verify that
argument forwarding and renaming are working correctly without executing the
underlying engine.

```{r}
rcall <- fit(threshold_stops(r1 = 20, min_size = 3L), tf_mini, dry_run = TRUE)
```

Convert the call to a list to inspect each element:

```{r}
str(as.list(rcall))
```

The `dry_run` flag itself is stripped from the call list so it never reaches
the engine.

Arguments passed via `...` to `fit()` are merged after the spec args, so
they can be used to override individual values for a single call without
modifying the spec:

```{r}
call_list_override <- as.list(
  fit(threshold_stops(), tf_mini, dry_run = TRUE, stop_id_col = "my_stop")
)
call_list_override[["stop_id_col"]]
```


# Complete worked example

The steps below show how to add a new model.

1. Write a specification constructor. If a constructor already exists and only a new engine is added, this step can be omitted.
2. Optionally wrap the engine function to handle validation or argument translation
   before registration.
3. Register the fit function.


## Specification constructor

The constructor collects parameters into a named list and passes them to
`new_travelpath_spec()`. The first argument is the model name (used as a key in the registry),
`args` holds canonical parameters shared across engines, and `eng_args` holds engine-specific
parameters not exposed in the constructor signature.

```{r}
my_model <- function(
  mode   = "segmentation",
  engine = "my_engine",
  threshold = 0.5,
  max_gap   = 100L
) {
  args <- list(threshold = threshold, max_gap = max_gap)
  new_travelpath_spec(
    "my_model",
    args     = args,
    eng_args = list(),
    mode     = mode,
    engine   = engine
  )
}
```


## Wrapper function

The wrapper function must accept `data` as its first positional argument,
followed by the same parameter names used in `args`.
The `...` argument allows `fit()` to forward additional arguments.
Here engine-specific validation can be added.

```{r}
my_fit_fn <- function(data, threshold = 0.5, max_gap = 100L, ...) {
  checkmate::assert_class(data, "trackframe")
  checkmate::assert_numeric(threshold, lower = 0, finite = TRUE, any.missing = FALSE)
  checkmate::assert_integerish(max_gap, lower = 0, any.missing = FALSE)
  stop_id <- c(1, sample(c(0, 1), NROW(data) - 1, replace = TRUE))
  data[["rand_stop_id"]] <- cumsum(stop_id)
  return(data)
}
```

## Register fit function

Register the wrapper under the `(model, mode, engine)` triple. In a package designed to enhance
the functionality of the travelpaths framework this call belongs in `.onLoad()`
inside `R/zzz.R` so the registry is populated when the package is attached.

```{r}
register_fit_function("my_model", "segmentation", "my_engine", my_fit_fn)
```

## Inspect registered arguments

The `arguments()` generic extracts the formal arguments of the fit function
associated with a specification.  This is useful for verifying that the
registered function matches what the spec expects:

```{r}
arguments(my_model())
```

## Use the model

Once registered, the model can be used with the standard `fit()` interface.

```{r}
data("tf_mini")
spec <- my_model()
fitted <- fit(spec, tf_mini)
fitted
```


# Packaging guidelines

When adding models in an extension package rather than a standalone script,
keep these points in mind:

- **Register in `.onLoad()`**, not at the top level.  Place
  `register_fit_function()` and `register_argument_mapping()` calls in
  `R/zzz.R` inside `.onLoad()`.  This ensures the registry is populated
  whenever the namespace is loaded.

- **Guard optional engines** with `is_pkg_installed()`.  If the engine
  package is listed under `Suggests` rather than `Imports`, wrap the
  registration in a conditional so the package still loads when the engine
  is absent:

```{r eval=FALSE}
.onLoad <- function(libname, pkgname) {
  if (is_pkg_installed("infostop")) {
    register_fit_function(
      "my_model", "segmentation", "infostop",
      infostop::identify_stops
    )
  }
}
```

- **Startup messages** from `.onAttach()` automatically list all
  registered models, so users see which engines are available.
