---
title: "Travel Paths"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Travel Paths}
  %\VignetteEngine{knitr::knitr}
editor_options: 
  chunk_output_type: console
---


```{r knitr_settings, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", out.width = "100%", error = TRUE)
knitr::opts_chunk$set(eval = FALSE) # Keep for now since it takes to long to run the whole vignette.
if (isTRUE(as.logical(Sys.getenv("NO_TRAVELPATHS_VIGNETTE")))) {
  knitr::opts_chunk$set(eval = FALSE)
}
```


# Package **travelpaths**

```{r setup, message=FALSE}
library("travelpaths")
library("infostop")

data("path_move2", package = "trackframe")
```


The **travelpaths** pipline is devided into 3 steps,

1. **segment**
2. **group**
3. **predict**



## Infostop

### Segment with `threshold_stops()`

```{r threshold_stops_spec}
args(threshold_stops)
```

```{r threshold_stops}
ths_spec <- threshold_stops(
  distance_metric = "haversine",
  r1 = 20,
  min_size = 3
)

stops <- fit(ths_spec, path_move2)
head(stops)
```

```{r infomap_spec}
args(infomap)
```


### Group with `infomap()`

```{r infomap_sites}
im_spec <- infomap()
sites <- fit(im_spec, stops)
head(sites)
```


## Change point test
### Detecting change points with `change_point_permtest()`
```{r change_point_permtest_spec}
args(change_point_permtest)
```

```{r change_point_permtest}
data("cpttestdata", package = "cpt")
cpt_spec <- change_point_permtest()
```

Inspect the arguments of the change point test specification:
```{r cpt_args}
arguments(cpt_spec)
```
We might should split it into `args` and `eng_args` sections.
Also we have to consider the mapping.

```{r cpt_update}
cpt_spec <- update(cpt_spec, q = 5, n = 10000, seed = 0)
```

```{r cpt_fit}
cpt_tf <- fit(cpt_spec, cpttestdata)
head(cpt_tf)
```


## Smoove

```{r smoove_kestrel_data}
library(smoove)
data(Kestrel)
k <- Kestrel[3730:4150, ]
head(k)
```


```{r scan_track_example, eval=FALSE}
# with(k, scan_track(x=X, y=Y, time = timestamp))
```


```{r create_trackframe}
library(trackframe)
data <- as.trackframe(k,
  easting_col = "X",
  northing_col = "Y",
  crs = NA)
```


Inspect the arguments of the change point detection by cvm
```{r cvm_change_points_spec}
args(cvm_change_points)
```

Specify the model spec
```{r cvm_cpt_spec}
cpt_spec <- cvm_change_points(
  windowsize = 50,
  windowstep = 5,
  model = "RACVM",
  time_unit = "secs"
)
```

Estimate the change points
```{r cvm_cpt_fit}
cpt_fit <- fit(cpt_spec, data)
head(cpt_fit)
```

Specify the phases object
```{r cvm_phases_spec}
args(cvm_phases)

phases_spec <- cvm_phases()
```

Fit the phases (grouping)
```{r cvm_phases_fit}
phases <- fit(phases_spec, cpt_fit)
head(phases)
```

